---
title: "Honolulu Gragons & Makiki Excelsior's Team Stats"
format:
  html:
    theme:
      - brand
      - minty
    css: styles-2025.css
    toc: true
    toc-location: right
execute:          
  echo: false     
  message: false
  warning: false
---

```{r}
library(lubridate)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(tidyverse)
  df <- read.csv("data/baseball _records.csv") %>%
  mutate(
  date = as.Date(date),
  year = lubridate::year(date)
)

df_2025 <- df %>% filter(year == 2025)



#df %>% filter(year(date) == 2025, sport == 'b') %>% 
  #summarise(unique_games = n_distinct(game_id))


```

### ⚾︎Table Explanation⚾︎

**game**: Number of games you played　（出場試合数）

**AB**: At-bats （打席数）

**HIT**: Total hits　（安打）

**AVG**: Batting Average: HIT/AB　（打率　＝　安打/打席数）

**RBI**: Runs batted in （打点）

**Single**: Total 1B （単打）

**Doubles**: Total 2B　（二塁打）

**Triple**: Total 3B　（三塁打）

**HR**: Home runs　（ホームラン）

**BB**: Walks （四球）

**HBP**: Hit by pitch　（死球）

**SF**: Sacrifice flies　（犠牲フライ）

**TB**: Total Bases = 1×Single + 2×Double + 3×Triple + 4×HR　（塁打 ＝ 1**×**単打 + 2×二塁打 + 3×三塁打 + 4×ホームラン）

**SLG**: Slugging percentage = TB / AB.　（長打率　＝　塁打/打席数）

**OBP**: On-base percentage = (HIT + BB + HBP) / (AB + BB + HBP + SF).（出塁率　＝　(安打数 + 四球数 + 死球数) ÷ (打数 + 四球数 + 死球数 + 犠飛数)）

**OPS**: On-base percentage + slugging = OBP + SLG.　（OPS ＝　出塁率　＋　長打率）　

**“Players highlighted are qualified players (met the minimum AB).”**

※ハイライトされている選手は、規定打席（または規定打数）に到達しています。

#  ⚾︎Total Stats(Baseball & Softball)⚾︎

1.  <div>

    ```{r}

    ####### Data cleaning ########
    # hit_type (1B, !B, and 1b)
    df <- df_2025 %>% 
      mutate(
        hit_type = str_trim(hit_type),
        hit_type = na_if(hit_type, ""), # "" -> NA
        hit_type = str_to_upper(hit_type), # 1b -> 1B
        hit_type = recode(hit_type, '!B' = "1B") # str_to_upper changes !b to !B.
      )
    #sort(unique(df$hit_type)) #Check datas were cleaned correctly


    # Total Hit Records
    ## Total At Bat
    total_at_bat_b <- df_2025 %>% 
      filter(event %in% c("HIT",'OUT', 'E', 'K')) %>% 
      count(player_name, name = "AB") %>% 
      arrange(desc(AB))

    ## Total rbi
    total_rbi_b <- df_2025 %>% 
      group_by(player_name) %>% 
      summarise(rbi_b = sum(rbi, na.rm = TRUE)) %>% 
      arrange(desc(rbi_b))

    ## Total 1B 
    total_one_base_b <- df_2025 %>% 
      filter(hit_type == '1B') %>% 
      mutate(one_b_b = hit_type == '1B', '1b', '!b') %>% 
      group_by(player_name) %>% 
      summarise(single = sum(one_b_b)) %>% 
      arrange(desc(single))

    ## Total 2B 
    total_two_base_b <- df_2025 %>% 
      filter(hit_type == '2B') %>% 
      mutate(two_b_b = hit_type == '2B') %>% 
      group_by(player_name) %>% 
      summarise(double = sum(two_b_b, na.rm = TRUE)) %>% 
      arrange(desc(double))

    #total_two_base_b

    ## Total 3B 
    total_three_base_b <- df_2025%>% 
      filter(hit_type == '3B') %>% 
      mutate(three_b_b = hit_type == '3B') %>% 
      group_by(player_name) %>% 
      summarise(triple = sum(three_b_b, na.rm = TRUE)) %>% 
      arrange(desc(triple))

    #total_three_base_b

    ## Total HR 
    total_hr_base_b <- df_2025 %>% 
      filter(hit_type == 'HR') %>% 
      mutate(hr_b = hit_type == 'HR') %>% 
      group_by(player_name) %>% 
      summarise(hr = sum(hr_b, na.rm = TRUE)) %>% 
      arrange(desc(hr))

    ## Count the game 
    total_games <- df_2025 %>% 
      group_by(player_name) %>% 
      summarise(game = n_distinct(game_id), .groups = "drop") %>% 
      arrange(desc(game))

    #total_games

    #total_hr_base_b

    ## Totla stats for both baseball and softball 
    at_bat <- df_2025 %>% 
      mutate(rem_sac = (event == "OUT") & !is.na(rbi) & (rbi >= 1) & (field_pos %in% c(7,8,9)), # Remove scrify fly 
             total_ab = event %in% c("HIT",'OUT', 'E', 'K') & !rem_sac, # Remove BB, and Db and Sacrify fly
             hit = event == 'HIT',
             one_base = hit_type == '1B',
             two_base = hit_type == '2B',
             three_base = hit_type == '3B',
             hr = hit_type == 'HR',
             bb = event == 'BB',
             db = event == 'DB') %>%
      group_by(player_name) %>% 
      summarise(
        game = n_distinct(game_id), .groups = "drop",
        AB = sum(total_ab, na.rm = TRUE),
        HIT = sum(hit, na.rm = TRUE),
        AVG = HIT/AB,
        RBI = sum(rbi, na.rm = TRUE),
        Single = sum(one_base, na.rm = TRUE),
        Double = sum(two_base, na.rm = TRUE), 
        Triple = sum(three_base, na.rm = TRUE),
        HR = sum(hr, na.rm = TRUE),
        BB = sum(bb, na.rm = TRUE),
        HBP = sum(db, na.rm = TRUE),
        SF = sum(rem_sac, na.rm = TRUE),
        TB = Single*1 + Double*2 + Triple*3 + HR*4,
        SLG = TB/AB, #SLG
        OBP = (HIT + BB + HBP) / (AB + BB + SF + HBP),
        OPS = SLG + OBP
        
        ) %>% 
      arrange(desc(AB))

    #at_bat

    ####### hightlight the quarify player #########
    qualify_players <- at_bat %>% 
      mutate(qualify_player = ceiling(n_distinct(df$game_id) * 2.3), # make sure the qualify player
             hightlight = AB >= qualify_player) # hightlight qualify player in kable

    ## Remove TB and highlight column in kable
    game_stats <- qualify_players %>% 
      select(!c(hightlight, qualify_player))

    ### Display the 3 decimal on AVG, OPS, OBP, SLG.
    three_decimal <- function(x) {
      out <- ifelse(is.na(x),"",sprintf("%.3f", x))
      out <- ifelse(out != "" & 1 > x & x >= 0, sub("^0", "", out), out) # if the number grater than 1, keep them. 
      out
    } 

    ## adding to kable
    game_stats_total <- game_stats %>% 
      mutate(across(c(AVG, SLG, OBP, OPS), three_decimal))





    ###### Create a Table ######
    kable(game_stats_total,
          caption = "30 baseball and softball games in 2025",
          booktabs = TRUE) %>%
      kableExtra::row_spec(which(qualify_players$hightlight), background = "lightcyan1") %>%
      kable_styling(bootstrap_options = c("condensed", "hover", "bordered")) %>%  # <- removed "basic"
      row_spec(0, bold = TRUE, color = "black", background = "white") %>%
      column_spec(1, bold = TRUE, color = "white", background = "blue1") %>%
      footnote(general = "This order is based on numbers of total at bat")

    ```

    </div>

# ⚾️Baseball Total Stats⚾️

```{r}
## Tatoal Stats for baseball
at_bat_base <- df_2025 %>% 
  filter(sport == 'b') %>% 
  mutate(rem_sac = (event == "OUT") & !is.na(rbi) & (rbi >= 1) & (field_pos %in% c(7,8,9)), # Remove scrify fly 
         total_ab = event %in% c("HIT",'OUT', 'E', 'K') & !rem_sac, # Remove BB, and Db and Sacrify fly
         hit = event == 'HIT',
         one_base = hit_type == '1B',
         two_base = hit_type == '2B',
         three_base = hit_type == '3B',
         hr = hit_type == 'HR',
         bb = event == 'BB',
         db = event == 'DB') %>%
  group_by(player_name) %>% 
  summarise(
    game = n_distinct(game_id), .groups = "drop",
    AB = sum(total_ab, na.rm = TRUE),
    HIT = sum(hit, na.rm = TRUE),
    AVG = HIT/AB,
    RBI = sum(rbi, na.rm = TRUE),
    Single = sum(one_base, na.rm = TRUE),
    Double = sum(two_base, na.rm = TRUE), 
    Triple = sum(three_base, na.rm = TRUE),
    HR = sum(hr, na.rm = TRUE),
    BB = sum(bb, na.rm = TRUE),
    HBP = sum(db, na.rm = TRUE),
    SF = sum(rem_sac, na.rm = TRUE),
    TB = Single*1 + Double*2 + Triple*3 + HR*4,
    SLG = TB/AB, #SLG
    OBP = (HIT + BB + HBP) / (AB + BB + SF + HBP),
    OPS = SLG + OBP
    ) %>% 
  arrange(desc(AB))

#at_bat_base


## Count the baseball game for makr sure the quarify player
games_b_2025 <- df_2025 %>% 
  filter(sport == 'b') %>% 
  summarise(games = n_distinct(game_id)) %>% 
  pull(games)

####### hightlight the quarify player #########
qualify_players_base <- at_bat_base %>%
  mutate(
    qualify_AB = ceiling(games_b_2025 * 2.0),
    highlight = AB >= qualify_AB
  ) # hightlight qualify player in kable




### Display the 3 decimal on AVG, OPS, OBP, SLG.
three_decimal <- function(x) {
  out <- ifelse(is.na(x),"",sprintf("%.3f", x))
  out <- ifelse(out != "" & 1 > x & x >= 0, sub("^0", "", out), out) # if the number grater than 1, keep them. 
  out
} 

## adding to kable 
game_stats_total_base <- qualify_players_base %>% 
  mutate(across(c(AVG, SLG, OBP, OPS), three_decimal))


# rows to highlight 
hl_rows <- which(game_stats_total_base$highlight)

# table to display (remove helper columns)
game_stats_total_base_show <- game_stats_total_base %>%
  select(-highlight, -qualify_AB)

###### Create a Table ######

kable(game_stats_total_base_show, caption = "11 baaseball games in 2025", booktabs = TRUE) %>%
  kableExtra::row_spec(hl_rows, background = "lightcyan1") %>% 
  kable_styling(
    bootstrap_options = c("condensed", "hover", "bordered")
  ) %>% 
  row_spec(0, bold = TRUE, color = "black", background = "white") %>%   column_spec(1, bold = TRUE, color = "white", background = "turquoise") %>% 
  footnote(general = "This order is based on numbers of total at bat")



```

# 🥎Softball Total Stats🥎

```{r}
## Tatoal Stats for baseball
at_bat_soft <- df_2025 %>% 
  filter(sport == 's') %>% 
  mutate(rem_sac = (event == "OUT") & !is.na(rbi) & (rbi >= 1) & (field_pos %in% c(7,8,9)), # Remove scrify fly 
         total_ab = event %in% c("HIT",'OUT', 'E', 'K') & !rem_sac, # Remove BB, and Db and Sacrify fly
         hit = event == 'HIT',
         one_base = hit_type == '1B',
         two_base = hit_type == '2B',
         three_base = hit_type == '3B',
         hr = hit_type == 'HR',
         bb = event == 'BB',
         db = event == 'DB') %>%
  group_by(player_name) %>% 
  summarise(
    game = n_distinct(game_id), .groups = "drop",
    AB = sum(total_ab, na.rm = TRUE),
    HIT = sum(hit, na.rm = TRUE),
    AVG = HIT/AB,
    RBI = sum(rbi, na.rm = TRUE),
    Single = sum(one_base, na.rm = TRUE),
    Double = sum(two_base, na.rm = TRUE), 
    Triple = sum(three_base, na.rm = TRUE),
    HR = sum(hr, na.rm = TRUE),
    BB = sum(bb, na.rm = TRUE),
    HBP = sum(db, na.rm = TRUE),
    SF = sum(rem_sac, na.rm = TRUE),
    TB = Single*1 + Double*2 + Triple*3 + HR*4,
    SLG = TB/AB, #SLG
    OBP = (HIT + BB + HBP) / (AB + BB + SF + HBP),
    OPS = SLG + OBP
    ) %>% 
  arrange(desc(AB))

#at_bat_soft


## Count the baseball game for makr sure the quarify player
games_s_2025 <- df_2025 %>% 
  filter(sport == 's') %>% 
  summarise(games = n_distinct(game_id)) %>% 
  pull(games)

####### hightlight the quarify player #########
qualify_players_soft <- at_bat_soft %>%
  mutate(
    qualify_AB = ceiling(games_s_2025 * 2.2),
    highlight = AB >= qualify_AB
  ) # hightlight qualify player in kable


### Display the 3 decimal on AVG, OPS, OBP, SLG.
three_decimal <- function(x) {
  out <- ifelse(is.na(x),"",sprintf("%.3f", x))
  out <- ifelse(out != "" & 1 > x & x >= 0, sub("^0", "", out), out) # if the number grater than 1, keep them. 
  out
} 


## adding to kable
game_stats_total_soft <- qualify_players_soft %>% 
  mutate(across(c(AVG, SLG, OBP, OPS), three_decimal))

# rows to highlight 
hl_rows <- which(game_stats_total_soft$highlight)

# table to display (remove helper columns)
game_stats_total_soft_show <- game_stats_total_soft %>%
  select(-highlight, -qualify_AB)



###### Create a Table ######
kable(game_stats_total_soft_show, caption = "19 softball games in 2025", booktabs = TRUE) %>%
  kableExtra::row_spec(hl_rows, background = "lightcyan1") %>% 
  kable_styling(
    bootstrap_options = c("condensed", "hover", "bordered")
  ) %>% 
  row_spec(0, bold = TRUE, color = "black", background = "white") %>%   column_spec(1, bold = TRUE, color = "black", background = "khaki") %>% 
  footnote(general = "This order is based on numbers of total at bat")
          
```

##  🥇🥈🥉Ranking for some categories

'\_s' refers to 'softball'.

'\_b' refers to 'baseball'.

```{r}
### makeing the ranking
make_ranking <- function(df_in, stat_col, n = 10, include_tie = FALSE) {
  # df_in: your table (ex: game_stats)
  # stat_col: column name as string (ex: "HIT")
  # n: how many players to show
  # include_ties_at_cutoff:
  #   FALSE -> always exactly n players
  #   TRUE  -> include everyone tied with the nth player (may show > n players)
  x <- df_in %>% 
    select(player_name, value = all_of(stat_col)) %>% 
    filter(!is.na(value) & value != 0) %>% 
    arrange(desc(value)) %>% 
    mutate(
      rank_num = dense_rank(desc(value)), # show the ranking numbers
      rank = paste0(rank_num, ".")
    )
  
    if (include_tie) {
      x <- x %>% filter(rank_num <= n) # include tied players at the cutoff
    } else {
      x <- x %>% slice_head(n=n) # display exactly n players
    }
  
  x %>% select(rank, player_name, value) %>% 
    rename(!!stat_col := value)
}



#######################################################################

make_player_summary <- function(df_in) {
  df_in %>%
    mutate(
      hit    = event == "HIT",
      single = hit & hit_type == "1B",
      double = hit & hit_type == "2B",
      triple = hit & hit_type == "3B",
      hr     = hit & hit_type == "HR"
    ) %>%
    group_by(player_name, sport) %>%
    summarise(
      games  = n_distinct(game_id),
      HIT    = sum(hit, na.rm = TRUE),
      Single = sum(single, na.rm = TRUE),
      Double = sum(double, na.rm = TRUE),
      Triple = sum(triple, na.rm = TRUE),
      HR     = sum(hr, na.rm = TRUE),
      RBI    = sum(rbi, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    pivot_wider(
      names_from  = sport,
      values_from = c(games, HIT, Single, Double, Triple, HR, RBI),
      names_sep   = "_",
      values_fill = 0
    ) %>%
    mutate(
      games_total  = games_b + games_s,
      HIT_total    = HIT_b + HIT_s,
      Single_total = Single_b + Single_s,
      Double_total = Double_b + Double_s,
      Triple_total = Triple_b + Triple_s,
      HR_total     = HR_b + HR_s,
      RBI_total = RBI_b + RBI_s
    ) %>%
    select(
      player_name,
      games_total, games_s, games_b,
      HIT_total, HIT_s, HIT_b,
      Single_total, Single_s, Single_b,
      Double_total, Double_s, Double_b,
      Triple_total, Triple_s, Triple_b,
      HR_total, HR_s, HR_b,
      RBI_total, RBI_b, RBI_s
    ) %>%
    arrange(desc(HIT_total))
}

# Assign the function
summary_2025 <- make_player_summary(df_2025)


## Ranking for Total HIT ##
top10_hit <- make_ranking(summary_2025, stat_col = "HIT_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, HIT_s, HIT_b),
    by = "player_name"
  )

# Create a table for Hit ranking table 
kable(
  top10_hit,
  caption = "Total Hit Ranking Table"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "Top 10 players")


## Single Ranking##
top10_single <- make_ranking(summary_2025, stat_col = "Single_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, Single_s, Single_b),
    by = "player_name"
  )

kable(top10_single, caption = "Total 10 Single Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "Top 10 players")

## double Ranking##
top10_double <- make_ranking(summary_2025, stat_col = "Double_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, Double_s, Double_b),
    by = "player_name"
  )

kable(top10_double, caption = "Total 10 Double Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "Top 10 players")


## Triple Ranking##
top10_triple <- make_ranking(summary_2025, stat_col = "Triple_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, Triple_s, Triple_b),
    by = "player_name"
  )

kable(top10_triple, caption = "Total 10 Triple Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "Top 10 players")


## HR Ranking ##
top10_hr <- make_ranking(summary_2025, stat_col = "HR_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, HR_s, HR_b),
    by = "player_name"
  )

kable(top10_hr, caption = "Top 10 HR Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "Top 10 players")


# Ranking RBI
top10_rbi  <- make_ranking(summary_2025, stat_col = "RBI_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, RBI_s, RBI_b),
    by = "player_name"
  )

kable(top10_rbi, caption = "Top 10 RBI Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "Top 10 players")


```

```{r}
# Games Played
top10_game <- make_ranking(summary_2025, stat_col = "games_total",n = 10, include_tie = TRUE)  %>% 
  left_join(
    summary_2025 %>% select(player_name, games_s, games_b),
    by = "player_name"
  )

kable(top10_game, caption = "Top 10 Players by Games Played") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  column_spec(1, bold = TRUE, color = "cornflowerblue", background = "lavenderblush") %>% 
  footnote(general = "30 games in 2025")
```

#### 30 games in 2025 (19 softball, 11 baseball)

\
