---
title: "Honolulu Dragons Team Records"
echo: false
message: false
warning: false
---

```{r}
df <- read.csv("data/baseball _records.csv")
```

# Total Stats

### Table Explanation

**game**: Number of games you played　（出場試合数）

**AB**: At-bats （打席数）

**HIT**: Total hits　（安打）

**AVG**: Batting Average: HIT/AB　（打率　＝　安打/打席数）

**RBI**: Runs batted in （打点）

**Single**: Total 1B （単打）

**Doubles**: Total 2B　（二塁打）

**Triple**: Total 3B　（三塁打）

**HR**: Home runs　（ホームラン）

**BB**: Walks （四球）

**HBP**: Hit by pitch　（死球）

**SF**: Sacrifice flies　（犠牲フライ）

**TB**: Total Bases = 1×Single + 2×Double + 3×Triple + 4×HR　（塁打 ＝ 1**×**単打 + 2×二塁打 + 3×三塁打 + 4×ホームラン）

**SLG**: Slugging percentage = TB / AB.　（長打率　＝　塁打/打席数）

**OBP**: On-base percentage = (HIT + BB + HBP) / (AB + BB + HBP + SF).（出塁率　＝　(安打数 + 四球数 + 死球数) ÷ (打数 + 四球数 + 死球数 + 犠飛数)）

**OPS**: On-base percentage + slugging = OBP + SLG.　（OPS ＝　出塁率　＋　長打率）　

**“Players highlighted in light blue are qualified players (met the minimum AB).”**

※水色でハイライトされている選手は、規定打席（または規定打数）に到達しています。

```{r}
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(tidyverse)
####### Data cleaning ########
# hit_type (1B, !B, and 1b)
df <- df %>% 
  mutate(
    hit_type = str_trim(hit_type),
    hit_type = na_if(hit_type, ""), # "" -> NA
    hit_type = str_to_upper(hit_type), # 1b -> 1B
    hit_type = recode(hit_type, '!B' = "1B") # str_to_upper changes !b to !B.
  )
#sort(unique(df$hit_type)) #Check datas were cleaned correctly


# Total Hit Records
## Total At Bat
total_at_bat_b <- df %>% 
  filter(event %in% c("HIT",'OUT', 'E', 'K')) %>% 
  count(player_name, name = "AB") %>% 
  arrange(desc(AB))

## Total rbi
total_rbi_b <- df %>% 
  group_by(player_name) %>% 
  summarise(rbi_b = sum(rbi, na.rm = TRUE)) %>% 
  arrange(desc(rbi_b))

## Total 1B 
total_one_base_b <- df %>% 
  filter(hit_type == '1B') %>% 
  mutate(one_b_b = hit_type == '1B', '1b', '!b') %>% 
  group_by(player_name) %>% 
  summarise(single = sum(one_b_b)) %>% 
  arrange(desc(single))

## Total 2B 
total_two_base_b <- df %>% 
  filter(hit_type == '2B') %>% 
  mutate(two_b_b = hit_type == '2B') %>% 
  group_by(player_name) %>% 
  summarise(double = sum(two_b_b, na.rm = TRUE)) %>% 
  arrange(desc(double))

#total_two_base_b

## Total 3B 
total_three_base_b <- df %>% 
  filter(hit_type == '3B') %>% 
  mutate(three_b_b = hit_type == '3B') %>% 
  group_by(player_name) %>% 
  summarise(triple = sum(three_b_b, na.rm = TRUE)) %>% 
  arrange(desc(triple))

#total_three_base_b

## Total HR 
total_hr_base_b <- df %>% 
  filter(hit_type == 'HR') %>% 
  mutate(hr_b = hit_type == 'HR') %>% 
  group_by(player_name) %>% 
  summarise(hr = sum(hr_b, na.rm = TRUE)) %>% 
  arrange(desc(hr))

## Count the game 
total_games <- df %>% 
  group_by(player_name) %>% 
  summarise(game = n_distinct(game_id), .groups = "drop") %>% 
  arrange(desc(game))

#total_games

#total_hr_base_b

## CALUCULATE THE OPS -> SLG + OBP
at_bat <- df %>% 
  mutate(rem_sac = (event == "OUT") & !is.na(rbi) & (rbi >= 1) & (field_pos %in% c(7,8,9)), # Remove scrify fly 
         total_ab = event %in% c("HIT",'OUT', 'E', 'K') & !rem_sac, # Remove BB, and Db and Sacrify fly
         hit = event == 'HIT',
         one_base = hit_type == '1B',
         two_base = hit_type == '2B',
         three_base = hit_type == '3B',
         hr = hit_type == 'HR',
         bb = event == 'BB',
         db = event == 'DB') %>%
  group_by(player_name) %>% 
  summarise(
    game = n_distinct(game_id), .groups = "drop",
    AB = sum(total_ab, na.rm = TRUE),
    HIT = sum(hit, na.rm = TRUE),
    AVG = HIT/AB,
    RBI = sum(rbi, na.rm = TRUE),
    Single = sum(one_base, na.rm = TRUE),
    Double = sum(two_base, na.rm = TRUE), 
    Triple = sum(three_base, na.rm = TRUE),
    HR = sum(hr, na.rm = TRUE),
    BB = sum(bb, na.rm = TRUE),
    HBP = sum(db, na.rm = TRUE),
    SF = sum(rem_sac, na.rm = TRUE),
    TB = Single*1 + Double*2 + Triple*3 + HR*4,
    SLG = TB/AB, #SLG
    OBP = (HIT + BB + HBP) / (AB + BB + SF + HBP),
    OPS = SLG + OBP
    
    ) %>% 
  arrange(desc(game))

#at_bat

####### hightlight the quarify player #########
qualify_players <- at_bat %>% 
  mutate(qualify_player = ceiling(n_distinct(df$game_id) * 2.3), # make sure the qualify player
         hightlight = AB >= qualify_player) # hightlight qualify player in kable

## Remove TB and highlight column in kable
game_stats <- qualify_players %>% 
  select(!c(hightlight, qualify_player))

### Display the 3 decimal on AVG, OPS, OBP, SLG.
three_decimal <- function(x) {
  out <- ifelse(is.na(x),"",sprintf("%.3f", x))
  out <- ifelse(out != "" & 1 > x & x >= 0, sub("^0", "", out), out) # if the number grater than 1, keep them. 
  out
} 

## adding to kable
game_stats_total <- game_stats %>% 
  mutate(across(c(AVG, SLG, OBP, OPS), three_decimal))





###### Create a Table ######
kable(game_stats_total, caption = "Total Stats for basedball (軟式）and softball", booktabs = TRUE) %>%
  kableExtra::row_spec(which(qualify_players$hightlight), background = "slategray1") %>% 
  kable_styling(
    bootstrap_options = c("condensed", "basic", "hover")
  ) %>% 
  row_spec(0, bold = TRUE, color = "white", background = "steelblue1") %>%   column_spec(1, bold = TRUE, color = "white", background = "steelblue2") %>% 
  footnote(general = "This order based on HIT")

```

## Ranking for some categories

```{r}

# Ranking for Total HIT
top10_hit <- game_stats_total %>% 
  select(player_name, HIT) %>% 
  arrange(desc(HIT)) %>% 
  slice_head(n = 10)


# Create a table for Hit ranking table 
kable(
  top10_hit,
  caption = "Total Hit Ranking Table"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "Top 10 players")

# Ranking single
top10_single <- game_stats_total %>% 
  select(player_name, Single) %>% 
  arrange(desc(Single)) %>% 
  slice_head(n = 10)

kable(top10_single, caption = "Total 10 Single Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "Top 10 players")

# Ranking single
top10_single <- game_stats_total %>% 
  select(player_name, Single) %>% 
  arrange(desc(Single)) %>% 
  slice_head(n = 10)

kable(top10_single, caption = "Total 10 Single Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "Top 10 players")

# Ranking double
top10_double <- game_stats_total %>% 
  select(player_name, Double) %>% 
  arrange(desc(Double)) %>% 
  slice_head(n = 11)

kable(top10_double, caption = "Total 10 Double Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "Top 10 players")

# Ranking triple
top10_triple <- game_stats_total %>% 
  select(player_name, Triple) %>% 
  arrange(desc(Triple)) %>% 
  slice_head(n = 5)

kable(top10_triple, caption = "Total 10 Triple Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "Top 10 players")


# Ranking HR
top10_hr <- game_stats_total %>% 
  select(player_name, HR) %>% 
  arrange(desc(HR)) %>% 
  slice_head(n = 13)

kable(top10_hr, caption = "Total 10 HR Ranking Table") %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "Top 10 players")
```

\

## Batting Average of baseball (軟式）

```{r}
batting_ave_b <- df %>% 
    filter(sport == 'b') %>% 
    mutate(
        total = event != "BB",
        hit = event == "HIT"
    ) %>% 
    group_by(player_name) %>% 
    summarise(
      AB = sum(total, na.rm = TRUE),
      H = sum(hit, na.rm = TRUE),
      AV = H/AB)  %>% 
  arrange(desc(H))

batting_ave_b
```

# Total Number of HIT

```{r}
#Total 
total_hit_both <- df %>% 
  filter(event == 'HIT') %>% 
  mutate(hit = event == "HIT") %>% 
  group_by(player_name) %>% 
  summarise(H = sum(hit, na.rm = TRUE))  %>% 
  arrange(desc(H))

total_hit_both

# Softball
total_hit_s <- df %>% 
  filter(event == 'HIT', sport == 's') %>% 
  mutate(hit = event == "HIT") %>% 
  group_by(player_name) %>% 
  summarise(H = sum(hit, na.rm = TRUE))  %>% 
  arrange(desc(H))

total_hit_s

# Baseball
total_hit_b <- df %>% 
  filter(event == 'HIT', sport == 'b') %>% 
  mutate(hit = event == "HIT") %>% 
  group_by(player_name) %>% 
  summarise(H = sum(hit, na.rm = TRUE))  %>% 
  arrange(desc(H))

total_hit_b

```

```{r}
#install.packages("kableExtra")

library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

# Total hits for BOTH
hits_both <- df %>%
  filter(event == "HIT") %>%
  count(player_name, name = "total hit")

# Total hits for SOFTBALL
hits_softball <- df %>%
  filter(event == "HIT", sport == "s") %>%
  count(player_name, name = "total_hit_softball")

# Total hits for BASEBALL
hits_baseball <- df %>%
  filter(event == "HIT", sport == "b") %>%
  count(player_name, name = "total_hit_baseball")

# Combine into one table (include players with 0 too)
hits_table <- df %>%
  distinct(player_name) %>%
  left_join(hits_both, by = "player_name") %>%
  left_join(hits_softball, by = "player_name") %>%
  left_join(hits_baseball, by = "player_name") %>%
  mutate(
    `total hit` = replace_na(`total hit`, 0),
    total_hit_softball = replace_na(total_hit_softball, 0),
    total_hit_baseball = replace_na(total_hit_baseball, 0)
  ) %>%
  arrange(desc(`total hit`))

hits_table


```

```{r}

kable(
  hits_table,
  caption = "Total Hits (Both / Softball / Baseball)"
) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed")
  ) %>% 
  row_spec(0, bold = TRUE, color = "azure", background = "#6495ED") %>% 
  column_spec(1, bold = TRUE, color = "darkcyan") %>% 
  footnote(general = "THis is a general footnote")
```

```{r}
df %>% 
  filter(event == 'HIT') %>% 
  select(player_name ,hit_type) %>% 
  count(player_name, name = 'HIT') %>% 
  arrange(desc(HIT))

```

\
